# IE - Master in Business Analytics and Big Data
Recommendation Systems

Building a Collaborative Filtering Engine

Author: George Farajalla

========================================================

# Initialization

Import libraries

```{r}
library(Hmisc)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(magrittr)
library(data.table)
```

Load the critics dataset.

```{r}
folder_path <- "C:/Users/PC/Desktop/IE/Recommendation Engines/Data" # change path here

critics <- fread(file.path(folder_path, "critics.csv"))
```

# 1) Pearson correlation coefficient 
Calculate the Pearson correlation coefficient between Sophia and Nuria

**Expected correlation between Sophia and Nuria**: 0.3651484
 
```{r fig.width=7, fig.height=6}

c <- as.data.frame(filter(critics, critics$User == "Sophia" | critics$User == "Nuria" )) # filter for only sophia and nuria
c$User <- NULL # remove user column
c <- t(c) # transpose data frame
colnames(c) <- c("Nuria", "Sophia")
expected <- cor(c[,1], c[,2], use="pairwise.complete.obs") # calculate correlation between them

sprintf("Expected correlation between Sophia and Nuria: %f", expected )

```

# 2) Compare two movie critics
Using the calculated Pearson correlation in the previous section. Compare and show the movie ratings for Sophia and Nuria on a scatter plot. How is this compared with the same ratings of Maria and Nerea on a scatter plot? What about Chris and Jim? What it means?

```{r} 

Sophia_and_Nuria <- as.data.frame(filter(critics, critics$User == "Sophia" | critics$User == "Nuria" ))
Sophia_and_Nuria$User <- NULL
rownames(Sophia_and_Nuria) <- c("Nuria", "Sophia")
Sophia_and_Nuria <- as.data.frame(t(as.matrix(Sophia_and_Nuria)))

graph1 <- ggplot(data = Sophia_and_Nuria, aes(x = Sophia_and_Nuria$Sophia, y = Sophia_and_Nuria$Nuria)) + geom_point() + geom_smooth(method = "lm", se = F)
graph1

## Both users are slightly postively correlated in terms of ratings and movies

Maria_and_Nerea <- as.data.frame(filter(critics, critics$User == "Maria" | critics$User == "Nerea" ))
Maria_and_Nerea$User <- NULL
rownames(Maria_and_Nerea) <- c("Maria", "Nerea")

Maria_and_Nerea <- as.data.frame(t(as.matrix(Maria_and_Nerea)))

graph2 <- ggplot(data = Maria_and_Nerea, aes(x = Maria_and_Nerea$Maria, y = Maria_and_Nerea$Nerea)) + geom_point() + geom_smooth(method = "lm", se = F)

## Both users are negatively correlated in terms of ratings and movies

Chris_and_Jim <- as.data.frame(filter(critics, critics$User == "Chris" | critics$User == "Jim" ))
Chris_and_Jim$User <- NULL
rownames(Chris_and_Jim) <- c("Chris", "Jim")
Chris_and_Jim <- as.data.frame(t(as.matrix(Chris_and_Jim)))

graph3 <- ggplot(data = Chris_and_Jim, aes(x = Chris_and_Jim$Jim, y = Chris_and_Jim$Chris)) + geom_point() + geom_smooth(method = "lm", se = F)

## Both users are strongly postively correlated in terms of ratings and movies
```

# 3) Top Recommendations 
Return the top-5 recommended movies for Sophia Use the weighted average of every other critic's rankings.

**Expected result**:
3.7917013044215, 'The Matrix'
3.50776533175371, 'Forrest Gump'     
3.33118834864677, 'The Sixth Sense' 
3.11491825315719, 'Shakespeare in Love'
2.9124513228665, 'Blade Runner'

```{r fig.width=7, fig.height=6}

## the objective here is to apply the forumla we learned in class
## we need to obtain the top recommended movies for sophia by similar
## users which is why we need to use a weighted average.
## we need the correlations between sophia and the other users as well 
## as the average ratings of each user. 

c <- as.data.frame(critics[,-1])

ct <- t(c)
colnames(ct)<- t(critics[,1])
rownames(c)<-t(critics[,1])

ct <- as.data.frame(ct) ## transform to data frame, makes it easier to work with

pos <- which(colnames(ct) == "Sophia") ## Sophia's position

## average rating of Sophia, first element of the formula
Sophia_avg <- mean(ct$Sophia, na.rm = T)

## calculate correlations between sophia and all other users
corr_df <- c()

for (i in 1:ncol(ct)){
  
  to_store <- cor(ct[,i], ct[, pos], use="pairwise.complete.obs", method = "pearson")
  
  corr_df[i] <- to_store
  
  
}

## weight w
sum_cor <- sum(abs(corr_df)) - 1 ## remove the 1 contribution from the correlation of Sophia with herself


final <- Sophia_avg + ## scaling difference          # multiply by correlations
  (colSums((c[-pos,] - rowMeans(c[-pos,],na.rm = T)) * corr_df[-pos], na.rm = T) / sum_cor) # divide by sum 

ns <- c(is.na(ct[,pos])) ## movies sophia hasn't seen

head(sort(final[ns], decreasing = T))[1:5]



```

# 4) Top similar critics
Return the top 5 critics that are most similar to Sophia. 

**Expected result**:
0.9449112, 'Rachel'
0.5976143, 'Ana' 
0.5789794, 'Oriol'
0.4925922, 'Maria'
0.4273247, 'Carles'

```{r fig.width=7, fig.height=6}

ct <- dcast(melt(critics, id.vars = "User"), variable ~ User) ## transpose
ct <- as.data.frame(ct) ## convert to data frame
critics_names <- colnames(critics)
critics_names <- critics_names[2:21]
ct$variable <- NULL
rownames(ct) <- critics_names ## set col names

# start from 2 since number 1 is sophia but we disregard since it is correlation with itself
# end at 6 to get actual top 5
sort(sapply(ct, function(x){cor(ct$Sophia, x, use = "pairwise.complete.obs")}), decreasing = T)[2:6]


```